Difference in 2019 Canadian Federal Election if Everyone had Voted

### 2020-12-19 clean data scripts/clean_census_data.Rmd
fixed post stratification result

```{r}
#Create cell counts:
cell_counts <- census %>% group_by(province, sex, as.factor(age), employed, bornin_canada, edu) %>% count() %>% rename(age = "as.factor(age)")
#Create a .csv file
write.csv(cell_counts, "census_cleaned.csv")
```

>>>>

```{r}
#Create cell counts:
census <- 
  census %>%
  count(sex, edu, employed, bornin_canada, province, age) %>%
  group_by(sex, edu, employed, bornin_canada, province, age)

#Create a .csv file
write.csv(census, "census_cleaned.csv")
```

### 2020-12-19 output/paper.Rmd
fixed post stratification result

```{r, echo=FALSE}
#create a bar plot of sex categorized by parties
sex <-table(survey$votechoice, survey$sex)
barplot(sex, beside=TRUE, main="Figure 1: Respondent Vote Choice in 2019 CES by Sex", xlab= "Sex", ylab= "Number of Respondents", col = c("red", "blue"))
legend("bottom", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8, bg = "white")
```

@@

```{r, echo=FALSE}
#create a bar plot of college and above education categorized by parties
education <-table(survey$votechoice, survey$edu)
barplot(education, beside=TRUE, main="Figure 2: Respondent Vote Choice in 2019 CES by Education", xlab= "Education", ylab= "Number of Respondents", col = c("red", "blue"))
legend("topleft", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8)
```

@@

```{r, echo=FALSE}
#create a bar plot of employed for pay categorized by parties
employment <-table(survey$votechoice, survey$employed)
barplot(employment, beside=TRUE, main="Figure 3: Respondent Vote Choice in 2019 CES by Employment", xlab= "Employment", ylab= "Number of Respondents", col = c("red", "blue"))
legend("bottom", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8, bg = "white")

@@

```{r, echo=FALSE}
#create a bar plot of provinces categorized by parties
par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin
province <-table(survey$votechoice, survey$province)
barplot(province, beside=TRUE, horiz=TRUE, main="Figure 5: Respondent Vote Choice in 2019 CES by Provinces", xlab= "Number of Respondents", col = c("red", "blue"), cex.names=0.6)
legend("bottomright", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8)
```

@@

```{r, echo=FALSE}
#create a bar plot of age groups categorized by parties
par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin
age <-table(survey$votechoice, survey$age)
barplot(age, beside=TRUE, horiz=TRUE, main="Figure 6: Respondent Vote Choice in 2019 CES by Age Groups", xlab= "Number of Respondents", col = c("red", "blue"))
legend("bottomright", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8)
```
Younger voters under 44 years old prefer the Liberal more. Respondents over this age prefer the Conservative more. However, an exception occurs that there is a tie for the 60 to 74 age group.

@@

A subset of the 2017 General Social Survey (GSS) on the Family dataset is obtained for this report, which includes selected demographics and retrieved to post-stratify. The subset is cleaned by firstly selecting interested variables that match with the selected variables from survey data, then renaming some variables to keep consistency with the survey data. The selected variables contain sex, education, employment, province, born in Canada, and age group, which are all the factors that can be used to determine the vote for the Liberal Party or the Conservative Party. Similar to survey data, for the census data, we also reconstruct many variables into binary or categorical variables to fit the model later. Figures 7-12 display the distribution of the six reconstructed variables from census data. Since the graphs of binary variables sex, education, and employment show close proportion, please refers to appendix part 2 for Figures 7-9.

```{r, echo=FALSE}
born_p <-table(post$bornin_canada)
pie(born_p, main="Figure 10: Born in Canada Proportion in 2017 GSS", col = c("red", "blue"), cex.main = 0.85)
```

Born in Canada is a binary variable where 1 indicates "yes", and 0 indicates "no". According to Figure 10, there are many more people who are born in Canada. However, refer to Figure 4, we know both groups show equal like for both parties. Then this large difference proportion may not have any strong effect on the vote outcome.

```{r, echo=FALSE}
province_p <-table(post$province)
pie(province_p, main="Figure 11: Province Proportion in 2017 GSS", col = rainbow(10), cex.main = 0.85)
```

By the province distribution plot, we can see people from Ontario and Quebec take up the most proportion(nearly half), then people from other provinces split the rest of the half proportion, and people from Prince Edward Island take up the least.

```{r, echo=FALSE}
age_p <-table(post$age)
pie(age_p, main="Figure 12: Age Proportion in 2017 GSS", col = rainbow(5), cex.main = 0.85)

As shown by the pie plot of age groups, people aged between 30 to 74 take up over 3/4 of the entire respondents and they are equally divided. Then the rest of the respondents come from the youngest people aged between 18 to 29, and the eldest people aged 75 and above, and they are also equally divided. Hence, the main component for the final result is the votes from the middle ages.

@@

Multilevel logistic regression model estimates by interpreting regression coefficients:
$$\log \left(\frac{\hat{p}_{ij}}{1-\hat{p}_{ij}}\right)=0.37+0.57 x_{ij}^{sexMale}-0.66 x_{ij}^{educationed}+0.01 x_{ij}^{employed}-0.24 x_{ij}^{bornCA}$$
$$-0.05 x_{ij}^{age30-44}+0.24 x_{ij}^{age45-59}-0.02 x_{ij}^{age60-74}+0.14 x_{ij}^{age75above}$$
```{r,echo=FALSE}
#Statistical summary
kable(summary(model)[["coefficients"]], digits=4, 
      caption="Summary of Model Estimates")
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
##together vote outcome
#together <- post %>%
#mutate(proportion_voting_for_conservative = estimate*n) %>%
#  summarise(voting_for_conservative = sum(proportion_voting_for_conservative)/sum(n))
#kable(together, digits=4, 
#      caption="Vote Outcome for the Conservative Party of 2019 Canadian Federal Election if Everyone had Voted")

##summarize by geography
#geography <- post %>%
#mutate(proportion_voting_for_conservative = estimate*n) %>%
#  group_by(state) %>%
#  summarise(voting_for_conservative = sum(proportion_voting_for_conservative)/sum(n))
#kable(geography, digits=4, 
#      caption="Vote Outcome for the Conservative Party of 2019 Canadian Federal Election if Everyone had Voted by Province")

```
Errors exist in the code for getting vote outcome. To be fixed.

@@

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#create a bar plot of sex categorized by parties
sex <-table(survey$votechoice, survey$sex)
barplot(sex, beside=TRUE, main="Figure 1: Respondent Vote Choice in 2019 CES by Sex", xlab= "Sex", ylab= "Number of Respondents", col = c("red", "blue"))
legend("bottom", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8, bg = "white")
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#create a bar plot of college and above education categorized by parties
education <-table(survey$votechoice, survey$edu)
barplot(education, beside=TRUE, main="Figure 2: Respondent Vote Choice in 2019 CES by Education", xlab= "Education", ylab= "Number of Respondents", col = c("red", "blue"))
legend("topleft", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8)
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#create a bar plot of employed for pay categorized by parties
employment <-table(survey$votechoice, survey$employed)
barplot(employment, beside=TRUE, main="Figure 3: Respondent Vote Choice in 2019 CES by Employment", xlab= "Employment", ylab= "Number of Respondents", col = c("red", "blue"))
legend("bottom", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8, bg = "white")

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#create a bar plot of provinces categorized by parties
par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin
province <-table(survey$votechoice, survey$province)
barplot(province, beside=TRUE, horiz=TRUE, main="Figure 5: Respondent Vote Choice in 2019 CES by Provinces", xlab= "Number of Respondents", col = c("red", "blue"), cex.names=0.6)
legend("bottomright", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8)
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#create a bar plot of age groups categorized by parties
par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin
age <-table(survey$votechoice, survey$age)
barplot(age, beside=TRUE, horiz=TRUE, main="Figure 6: Respondent Vote Choice in 2019 CES by Age Groups", xlab= "Number of Respondents", col = c("red", "blue"))
legend("bottomright", legend= c("Conservative Party", "Liberal Party"), fill = c("red", "blue"), cex=0.8)
```
Younger voters under 44 years old prefer the Liberal more. Respondents over this age prefer the Conservative more. However, an exception occurs that there is a tie for the 60 to 74 age group.

@@

A subset of the 2017 General Social Survey (GSS) on the Family dataset is obtained for this report, which includes selected demographics and retrieved to post-stratify. The subset is cleaned by firstly selecting interested variables that match with the selected variables from survey data, then renaming some variables to keep consistency with the survey data. The selected variables contain sex, education, employment, province, born in Canada, and age group, which are all the factors that can be used to determine the vote for the Liberal Party or the Conservative Party. Similar to survey data, for the census data, we also reconstruct many variables into binary or categorical variables to fit the model later. Figures 7-12 display an even distribution trend for all the six reconstructed variables from census data. Due to the length of the six pie plots for the six variables, please refers to appendix part 2 for Figures 7-10 for the binary variables sex, education, employment, and born in Canada.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
province_p <-table(post$province)
pie(province_p, main="Figure 11: Province Proportion in 2017 GSS", col = rainbow(10), cex.main = 0.85)

age_p <-table(post$age)
pie(age_p, main="Figure 12: Age Proportion in 2017 GSS", col = rainbow(5), cex.main = 0.85)
```

By the pie plot of province and age groups according to Figure 11 and 12, we can see people ten provinces and five age groups equally split the entire population in the census dataset, which means people from every provinces and every age groups are well represented in the final vote outcome result.

@@

Multilevel logistic regression model estimates by interpreting regression coefficients:
$$\log \left(\frac{\hat{p}_{ij}}{1-\hat{p}_{ij}}\right)=0.37+0.57 x_{ij}^{sexMale}-0.66 x_{ij}^{educationed}+0.01 x_{ij}^{employed}-0.24 x_{ij}^{bornCA}$$
$$-0.05 x_{ij}^{age30-44}+0.24 x_{ij}^{age45-59}-0.02 x_{ij}^{age60-74}+0.14 x_{ij}^{age75above}$$
```{r,echo=FALSE, message=FALSE, warning=FALSE}
#Statistical summary
kable(summary(model)[["coefficients"]], digits=4, 
      caption="Summary of Model Estimates")
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#together vote outcome
together <- post %>%
mutate(proportion_voting_for_conservative = estimate*n) %>%
  summarise(voting_for_conservative = sum(proportion_voting_for_conservative)/sum(n))
kable(together, digits=4, 
      caption="Vote Outcome for the Conservative Party of 2019 Canadian Federal Election if Everyone had Voted")

#summarize by geography
geography <- post %>%
mutate(proportion_voting_for_conservative = estimate*n) %>%
  group_by(province) %>%
  summarise(voting_for_conservative = sum(proportion_voting_for_conservative)/sum(n))
kable(geography, digits=4, 
      caption="Vote Outcome for the Conservative Party of 2019 Canadian Federal Election if Everyone had Voted by Province")

@@

born_p <-table(post$bornin_canada)
pie(born_p, main="Figure 10: Born in Canada Proportion in 2017 GSS", col = c("red", "blue"), cex.main = 0.85)
```
